An assessment of photometric redshift PDF performance in the context of LSST

### 2018-09-10 paper/tex/results.tex
add clarification on overtraining
rebase after GPz merge, edit text on z sim 1.4 range

Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.  The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution. All distributions are smoothed via kernel density estimation (KDE) with a common bandwidth chosen via Scott's rule \citep{Scott:1992} in order to minimize differences in small-scale features and make for a more uniform comparison between codes.  While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
%New text for the new figure
As expected, \textsc{TrainZ} is in excellent agreement with true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.  \textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution. \textsc{FlexZBoost} and \textsc{TPZ} also show very good agreement, with only slight departures, with an over/under-prediction in the high redshift tail of $\hat{N}(z)$ evident around $z\sim1.4$.  In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.
This is likely due to the $4000$ angstrom break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.  The relative dearth of strong features blue-ward of the 4000 angstrom break seems to cause problems for the redshift range that corresponds to the break passing between the two reddest filters bandpasses of a survey.
The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$, and is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.
%end of new text (and slightly modified text)
Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

@@

It is also of note that all three template-based codes show an excess in their stacked $p(z)$ at $z\,\sim\,1.3-1.4$.  This redshift range corresponds to the wavelengths where the $4000$ Angstrom break is passing between the borders of the $z$ and $y$ filters.  This strong break entering the gap between the two reddest filters can cause problems with redshift estimation of individual galaxies, as can be seen in the point-estimate photo-$z$'s shown in Figure~\ref{fig:pz_pointestimates}.  This is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.  The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.  However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.  Further study on this issue may provide a solution for codes that suffer from this shortcoming.

@@

\begin{table}  %%% DATA TABLE %%%
\centering
\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
\begin{tabular}{lr}
\hline
\bf Photo-$z$ Code & \bf CDE Loss \\
\hline
\textsc{ANNz2} 		& $-6.88$ \\
\textsc{BPZ} 		& $-7.82$ \\
%\textsc{Delight} 	& $-4.06$ \\
\textsc{Delight}    & $-8.33$\\
%\textsc{EAZY} 		& $-7.97$ \\
\textsc{EAZY}       & $-7.07$ \\
%\textsc{FlexZBoost} & $-11.51$ \\
\textsc{FlexZBoost} & $-10.60$\\
\textsc{GPz} 		& $-9.56$ \\
\textsc{LePhare} 	& $-1.66$ \\
\textsc{METAPhoR} 	& $-6.28$ \\
%\textsc{CMNN} 		& $-$ \\
\textsc{CMNN}         & $-10.43$ \\
\textsc{SkyNet} 	& $-7.89$ \\
\textsc{TPZ} 		& $-9.55$ \\
\hline
\trainz		& $-0.83$ \\
%\textsc{Frankenz}	& $-$  \\
%\hline
\end{tabular}
\end{table}

\begin{table}
\setlength{\tabcolsep}{2pt}
\begin{center}
\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
\begin{tabular}{lcc}
\hline
\hline
Root-Mean-Square-Error \\
(RMSE) statistics \\
\hline
Photo-z Code & PIT RMSE & N(z) RMSE\\
\hline
\textsc{ANNz2}      & 0.019 & 0.0054\\
\textsc{BPZ}        & 0.032 & 0.0050\\
\textsc{Delight}    & 0.111 & 0.0056\\
\textsc{EAZY}       & 0.054 & 0.0102\\
\textsc{FlexZBoost} & 0.021 & 0.0022\\
\textsc{GPz}        & 0.048 & 0.0042\\
\textsc{LePhare}    & 0.028 & 0.0062\\
\textsc{METAPhoR}   & 0.064 & 0.0081\\
\textsc{CMNN}         & 0.108 & 0.0009\\
\textsc{Skynet}     & 0.054 & 0.0144\\
\textsc{TPZ}        & 0.082 & 0.0031\\
\hline
\trainz		& 0.0025 & 0.0013\\
\end{tabular}
\end{center}
\end{table}

>>>>

Fig.~\ref{fig:nz} shows the stacked $\hat{N}(z)$ distribution compared to the true redshift distribution $N'(z)$ for all tested codes.  The red line indicates the summed $p(z)$ for each code, while the blue line shows the true redshift distribution. All distributions are smoothed via kernel density estimation (KDE) with a common bandwidth chosen via Scott's rule \citep{Scott:1992} in order to minimize differences in small-scale features and make for a more uniform comparison between codes.  While Scott's rule is used to display $N'(z)$ in the figure, all quantitative statistics are computed via the empirical CDF, and are thus unaffected by bandwidth/smoothing choice.
%New text for the new figure
As expected, \textsc{TrainZ} is in excellent agreement with the true redshift distribution: as the training sample is selected from the same underlying distribution as the test set, the redshift distributions are identical, up to Poisson fluctuations due to the finite number of sample galaxies.  \textsc{CMNN} is also in excellent agreement for similar reasons: with a representative training sample of galaxies spanning the colour-space, the sum of the colour-matched neighbour redshifts should return the true redshift distribution. \textsc{FlexZBoost} and \textsc{TPZ} also show very good agreement, with only slight departures, with an over/under-prediction in the high redshift tail of $\hat{N}(z)$ evident around $z\sim1.4$.  In fact, several of the other codes show an excess at $z \sim 1.4$, particularly the template-based codes \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  This is likely due to the $4000\ {\rm \AA}$ break passing through the gap between the $z$ and $y$ filters, resulting in a drastic change in $z-y$ colour for galaxies in this redshift range.  With a relative dearth of strong features blue-ward of the $4000\ {\rm \AA}$ break in most galaxy SEDs, the colour change in the two reddest filter bandpasses of a survey has a large influence on the redshift determination.  The $z\sim1.4$ feature is one of the most prominent sources of larger uncertainty in individual galaxy $p(z)$.
In our sample individual galaxy $p(z)$'s tend to be broader around $z\sim1.4$ and point estimates are more uncertain in this regime, as is readily seen in the point-estimate plots shown in Fig.~\ref{fig:pz_pointestimates} and described in the Appendix.  This feature is not unique to this dataset, it is a common occurrence in photo-$z$ estimation.  The fact that similar excesses appear in Figure~\ref{fig:nz} for \textsc{ANNz2} and \textsc{METAPhoR} shows that the effect is not limited to template-based codes.  However, the lack of such a feature in the other codes shows that it is possible to eliminate the degeneracies.  Further study on this issue may provide a solution for codes that suffer from this shortcoming.

%end of new text (and slightly modified text)
Two of the machine learning based codes, \textsc{ANNz2} and \textsc{METAPhoR}, appear to be over-trained, adding excess galaxy probability to the redshift peaks with the largest number of training galaxies, and missing probability in the troughs where training galaxies are of fewer number.
Given that our training data is drawn from the same galaxy population as the test set, and our data has prominent peaks in $N'(z)$, perhaps it is not unexpected that such overtraining occurs in some codes, though the fact that it does not occur in all training-based codes indicates that it may be due to specifics of the implementations of \textsc{ANNz2} and \textsc{METAPhoR}.
A more extensive training/validation set might allow for a better choice of smoothing parameters in individual codes that would avoid such overtraining for these particular codes.
%\aim{Move the text on Scott's Rule to Discussion?}
%Scott's rule \citep{Scott:1992} was used to determine the smoothing bandwidth of the true redshift sample: that is, a single smoothing scale was chosen to represent the true redshift sample for all twelve codes.
%As with the $p(z)$ values in Figure~\ref{fig:pitqq}, different levels of substructure are obvious for the different codes.
%While Scott's rule provides a relatively good general smoothing scale to represent the true $N'(z)$, there are smaller scale fluctuations: while \textsc{FlexZBoost} and \textsc{CMNN} appear somewhat discrepant in Fig.~\ref{fig:nz}, they are actually the two most accurate in terms of their quantitative measurements.
%\textsc{SkyNet} also shows excess small-scale structure, but also shows significant overestimate of the high redshift galaxy population.
\textsc{SkyNet} shows an obvious redshift bias, evident both visually in Figure~\ref{fig:nz} and in the first moment of N(z) listed in Table~\ref{tab:moments}, where it is clearly an outlier.  \textsc{SkyNet} employed a method where a random sample of training galaxies was chosen, but there was no test that the subset was completely representative of the overall redshift distribution.  Unlike the previous implementation of \textsc{SkyNet} in \citet{Bonnett:15}, no effort was made to add extra weight to more rare low and high redshift galaxies.  Either of these decisions could be the cause of the bias seen in our results.  Future runs of \textsc{SkyNet} will explore these implementation choices and their effects.

@@

@@

\begin{table}  %%% DATA TABLE %%%
\centering
\caption{CDE loss statistic for each photo-$z$ code.} \label{tab:cdeloss}
\begin{tabular}{lr}
\hline
\bf Photo-$z$ Code & \bf CDE Loss \\
\hline
\textsc{ANNz2} 		& $-6.88$ \\
\textsc{BPZ} 		& $-7.82$ \\
%\textsc{Delight} 	& $-4.06$ \\
\textsc{Delight}    & $-8.33$\\
%\textsc{EAZY} 		& $-7.97$ \\
\textsc{EAZY}       & $-7.07$ \\
%\textsc{FlexZBoost} & $-11.51$ \\
\textsc{FlexZBoost} & $-10.60$\\
\textsc{GPz} 		& $-9.93$ \\
\textsc{LePhare} 	& $-1.66$ \\
\textsc{METAPhoR} 	& $-6.28$ \\
%\textsc{CMNN} 		& $-$ \\
\textsc{CMNN}         & $-10.43$ \\
\textsc{SkyNet} 	& $-7.89$ \\
\textsc{TPZ} 		& $-9.55$ \\
\hline
\trainz		& $-0.83$ \\
%\textsc{Frankenz}	& $-$  \\
%\hline
\end{tabular}
\end{table}

\begin{table}
\setlength{\tabcolsep}{2pt}
\begin{center}
\caption{Root-Mean-Square-Error (RMSE) statistics for the twelve photo-z codes for both PIT and $\hat{N}(z)$ distributions.}\label{tab:rmse}
\begin{tabular}{lcc}
\hline
\hline
Root-Mean-Square-Error \\
(RMSE) statistics \\
\hline
Photo-z Code & PIT RMSE & N(z) RMSE\\
\hline
\textsc{ANNz2}      & 0.019 & 0.0054\\
\textsc{BPZ}        & 0.032 & 0.0050\\
\textsc{Delight}    & 0.111 & 0.0056\\
\textsc{EAZY}       & 0.054 & 0.0102\\
\textsc{FlexZBoost} & 0.021 & 0.0022\\
\textsc{GPz}        & 0.027 & 0.0042\\
\textsc{LePhare}    & 0.028 & 0.0062\\
\textsc{METAPhoR}   & 0.064 & 0.0081\\
\textsc{CMNN}         & 0.108 & 0.0009\\
\textsc{Skynet}     & 0.054 & 0.0144\\
\textsc{TPZ}        & 0.082 & 0.0031\\
\hline
\trainz		& 0.0025 & 0.0013\\
\end{tabular}
\end{center}
\end{table}
